Manuscript Figures

### 2018-06-07 Manuscript Figures.Rmd
Why does maize1 average lower FPKM than maize2

```{r extra4, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=4, fig.height=4, fig.align='center'}
# Average expression for genes in each subgenome, biased by sample size.  While other plots show that maize1 dominates maize2 more often in all tissue, maize2 
# for pollen and endosperm crown the average expression is higher in maize2.  Basically, when maize2 wins, it wins by more raw FPKMs.  If we account for all
# genes in the subgenome, not just the pairs, maize2 often has higher average expression.  This is probably due to many very low expressed genes in maize1.
data <- 
  homeologs.pairs %>%
  # subset(Maize1 != "" & Maize2 != "") %>%
  # select(Maize1, Maize2) %>%
  distinct() %>%
  select(Maize1, Maize2) %>%
  left_join(maize.walley.v4mapped.expression, by=c("Maize1"="geneID")) %>%
  rename(FPKM_avg1=FPKM_avg) %>%
  left_join(maize.walley.v4mapped.expression, by=c("Maize2"="geneID", "sample"="sample")) %>%
  rename(FPKM_avg2=FPKM_avg) %>%
  select(Maize1, Maize2, sample, FPKM_avg1, FPKM_avg2) %>%
  # subset(sample %in% c("B73 Mature Pollen")) %>%
  group_by(sample) %>%
  summarize(Mean1=mean(FPKM_avg1, na.rm=TRUE), Mean2=mean(FPKM_avg2, na.rm=TRUE))
data <- 
  data %>%
  gather(subgenome, FPKM, -1)

@@

>>>>

```{r extra4, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=4, fig.height=4, fig.align='center'}
# Average expression for genes in each subgenome, biased by sample size.  While other plots show that maize1 dominates maize2 more often in all tissue, maize2 
# for pollen and endosperm crown the average expression is higher in maize2.  Basically, when maize2 wins, it wins by more raw FPKMs.  If we account for all
# genes in the subgenome, not just the pairs, maize2 often has higher average expression.  This is probably due to many very low expressed genes in maize1.
data <- 
  homeologs.pairs %>%
  # subset(Maize1 != "" & Maize2 != "") %>% #only matching
  # subset(!(Maize1 %in% c("") & Maize2 %in% c(""))) %>% #only nonmatching
  distinct() %>%
  select(Maize1, Maize2) %>%
  left_join(maize.walley.v4mapped.expression, by=c("Maize1"="geneID")) %>%
  rename(FPKM_avg1=FPKM_avg) %>%
  left_join(maize.walley.v4mapped.expression, by=c("Maize2"="geneID", "sample"="sample")) %>%
  rename(FPKM_avg2=FPKM_avg) %>%
  select(Maize1, Maize2, sample, FPKM_avg1, FPKM_avg2) %>%
  # subset(sample %in% c("B73 Mature Pollen")) %>%
  group_by(sample) %>%
  summarize(Mean1=mean(FPKM_avg1, na.rm=TRUE), Mean2=mean(FPKM_avg2, na.rm=TRUE))
data <- 
  data %>%
  gather(subgenome, FPKM, -1)

@@

```{r extra5, echo=FALSE, warning=FALSE, message=FALSE, results='asis', fig.width=4, fig.height=4, fig.align='center'}
## In maize1, there are 3672 matching genes and 7927 nonmatching.  Since they tend towards 1 (despite cutoff at 1), more will pull average down.
data <-
  homeologs.pairs %>%
  subset(Maize2 %in% c("")) %>% #only nonmatching
  select(Maize1) %>%
  left_join(maize.walley.v4mapped.expression, by=c("Maize1"="geneID")) %>%
  select(Maize1, sample, FPKM_avg)
temp <-
  homeologs.pairs %>%
  subset(Maize1 != "" & Maize2 != "") %>% #only matching
  select(Maize1) %>%
  left_join(maize.walley.v4mapped.expression, by=c("Maize1"="geneID")) %>%
  select(Maize1, sample, FPKM_avg)

data$status <- "nonMatch"
temp$status <- "match"
data <-
  data %>%
  bind_rows(temp)

ggplot(data, aes(status,log10(FPKM_avg))) +
  geom_boxplot() +
  labs(y = "log10(FPKM)",
       x = "Subgenome",
       title = "Comparison of FPKM Expression between\nGenes in Subgenome 1 and Subgenome 2"
  )
```
